IIIDYT at IEST 2018: Implicit Emotion Classification With Deep Contextualized Word Representations

### 2018-09-01 paper/paper.tex
[paper] Improve tables, begin writing Error Analysis subsection

\documentclass[11pt,a4paper]{article}
\pdfoutput=1
\usepackage[hyperref]{emnlp2018}
% \usepackage[showframe=true]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{times}
\usepackage{latexsym}
\usepackage{lipsum}
\usepackage{url}
\usepackage{multirow}
\usepackage{changepage}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage{caption}
\usepackage{blindtext}
\usepackage[hang,flushmargin]{footmisc} % remove indentation from footnotes
% \hypersetup{draft}

@@

    \begin{tabular}{ll}

        \textbf{Original} & \textbf{Replacement} \\
        \hline
        \hline
        \texttt{\footnotesize[\#TRIGGERWORD\#]} & \texttt{\footnotesize\_\_TRIGGERWORD\_\_} \\
        \texttt{\footnotesize@USERNAME} & \texttt{\footnotesize\_\_USERNAME\_\_} \\
        \texttt{\footnotesize[NEWLINE]} & \texttt{\footnotesize\_\_NEWLINE\_\_} \\
        \texttt{\footnotesize http://url.removed} & \texttt{\footnotesize\_\_URL\_\_} \\

@@

\textbf{Dimensionalities}: By default the ELMo layer outputs a 1024-dim vector,
which we then feed to a BiLSTM with output size 2048, resulting in a 4096-dim
vector when concatenating forward and backward directions for each word
of the sequence\footnote{A BiLSTM is composed of two separate LSTMs that read
the input in opposite directions and whose outputs are concatenated at the
hidden dimension. This results in a vector double the dimension of the input
for each time step.}. After max-pooling the BiLSTM output over the sequence
dimension, we obtain a single 4096-dim vector corresponding to the tweet
representation. This representation is finally fed to a single-layer
fully-connected network with input size 4096, 512 hidden units, output size 6,
and a ReLU nonlinearity after the hidden layer. The output of the
dense layer is a 6-dim logit vector for each input example.

@@

        \textbf{Variation} & \textbf{Accuracy (\%)} & $\bf{\Delta}$\textbf{\%} \\
        \hline
        \hline
        Submitted          & \textbf{69.23}         & -                        \\
        \hline
        No emoji           & 68.36                  & - 0.87                   \\
        \hline
        No ELMo            & 65.52                  & - 3.71                   \\
        \hline
        Concat Pooling     & 68.47                  & - 0.76                   \\
        \hline
        LSTM hidden=4096   & 69.10                  & - 0.13                   \\
        LSTM hidden=1024   & 68.93                  & - 0.30                   \\
        LSTM hidden=512    & 68.43                  & - 0.80                   \\
        \hline
        POS emb dim=100    & 68.99                  & - 0.24                   \\
        POS emb dim=75     & 68.61                  & - 0.62                   \\
        POS emb dim=50     & 69.33                  & + 0.10                   \\
        POS emb dim=25     & 69.21                  & - 0.02                   \\
        \hline
        SGD optim lr=1     & 64.33                  & - 4.90                   \\
        SGD optim lr=0.1   & 66.11                  & - 3.12                   \\
        SGD optim lr=0.01  & 60.72                  & - 8.51                   \\
        SGD optim lr=0.001 & 30.49                  & - 38.74                  \\

@@

Using a greater BiLSTM hidden size did not help the model, probably because of
the reason mentioned early; the fully-connected layer was not big or deep enough
to exploit the additional information. Similarly, using a smaller hidden size
neither helped.

We found that using 50-dim part-of-speech embeddings slightly improved results,
which implies that better fine-tuning this hyperparameter, or using a better POS
tagger could yield an even better performance.

@@

    \caption{Dropout Ablation. }
    \vspace{-0.4cm}

@@

@@

        \textbf{} & \textbf{Present} & \textbf{Not Present} \\
        \hline
        \hline
        Emoji     & 4805 (76.6\%)    & 23952 (68.0\%)       \\
        Hashtags  & 2122 (70.5\%)    & 26635 (69.4\%)       \\

@@

Table \ref{table:emoji_fine_grained} shows the effect specific emoji have on
classification performance.

\begin{table}[!h]
    \centering
    \scriptsize
        \begin{tabular}{lr|rc|rc|r}

           \multirow{2}{*}{\textbf{Emoji alias}} & \multirow{2}{*}{\textbf{N}} & \multicolumn{2}{c|}{\textbf{emoji}} & \multicolumn{2}{c|}{\textbf{no-emoji}} &
            \multirow{2}{*}{$\bf{\Delta}$\textbf{\%}} \\
            \cline{3-6}
                                &     & \#   & \% & \# & \% & \\
          \hline
          \hline
          \texttt{mask}         & 163 & 154 & 94.48 & 134 & 82.21 & - 12.27 \\
          \texttt{two\_hearts}  & 87  & 81  & 93.10 & 77  & 88.51 & - 4.59  \\
          \texttt{heart\_eyes}  & 122 & 109 & 89.34 & 103 & 84.43 & - 4.91  \\
          \texttt{heart}        & 267 & 237 & 88.76 & 235 & 88.01 & - 0.75  \\
          \texttt{rage}         & 92  & 78  & 84.78 & 66  & 71.74 & - 13.04 \\
          \texttt{cry}          & 116 & 97  & 83.62 & 83  & 71.55 & - 12.07 \\
          \texttt{sob}          & 490 & 363 & 74.08 & 345 & 70.41 & - 3.67  \\
          \texttt{unamused}     & 167 & 121 & 72.46 & 116 & 69.46 & - 3.00  \\
          \texttt{weary}        & 204 & 140 & 68.63 & 139 & 68.14 & - 0.49  \\
          \texttt{joy}          & 978 & 649 & 66.36 & 629 & 64.31 & - 2.05  \\
          \texttt{sweat\_smile} & 111 & 73  & 65.77 & 75  & 67.57 & 1.80 \\
          \texttt{confused}     & 77  & 46  & 59.74 & 48  & 62.34 & 2.60 \\

@@

It is clear some emoji strongly contribute to improving prediction quality. The
most interesting ones are \texttt{mask}, \texttt{rage}, and \texttt{cry}, which
significantly increase accuracy. Further, contrary to intuition, the
\texttt{sob} emoji contributes less than \texttt{cry}, despite representing a
stronger emotion. This is probably due to \texttt{sob} being used for depicting
a wider spectrum of emotions.

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=\columnwidth]{images/acc_vs_tdp_variation.eps}
    \caption{Validation accuracy and macro F1 score vs. proportion of the training data.}
    \label{fig:data_amt_vs_acc}
\end{figure}

>>>>

\documentclass[11pt,a4paper]{article}
\pdfoutput=1
\usepackage[hyperref]{emnlp2018}
% \usepackage[showframe=true]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{times}
\usepackage{latexsym}
\usepackage{lipsum}
\usepackage{url}
\usepackage{multirow}
\usepackage{changepage}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage{caption}
\usepackage{blindtext}
\usepackage{booktabs}
\usepackage[hang,flushmargin]{footmisc} % remove indentation from footnotes
% \hypersetup{draft}

@@

    \begin{tabular}{ll}
        \toprule
        \textbf{Original} & \textbf{Replacement} \\
        \midrule
        \texttt{\footnotesize[\#TRIGGERWORD\#]} & \texttt{\footnotesize\_\_TRIGGERWORD\_\_} \\
        \texttt{\footnotesize@USERNAME} & \texttt{\footnotesize\_\_USERNAME\_\_} \\
        \texttt{\footnotesize[NEWLINE]} & \texttt{\footnotesize\_\_NEWLINE\_\_} \\
        \texttt{\footnotesize http://url.removed} & \texttt{\footnotesize\_\_URL\_\_} \\
        \bottomrule

@@

\textbf{Dimensionalities}: By default the ELMo layer outputs a 1024-dimensional
vector, which we then feed to a BiLSTM with output size 2048, resulting in a
4096-dimensional vector when concatenating forward and backward directions for
each word of the sequence\footnote{A BiLSTM is composed of two separate LSTMs
that read the input in opposite directions and whose outputs are concatenated at
the hidden dimension. This results in a vector double the dimension of the input
for each time step.}. After max-pooling the BiLSTM output over the sequence
dimension, we obtain a single 4096-dimensional vector corresponding to the tweet
representation. This representation is finally fed to a single-layer
fully-connected network with input size 4096, 512 hidden units, output size 6,
and a ReLU nonlinearity after the hidden layer. The output of the dense layer is
a 6-dimensional logit vector for each input example.

@@

        \toprule
        \textbf{Variation} & \textbf{Accuracy (\%)} & $\bf{\Delta}$\textbf{\%} \\
        \midrule[0.08em]
        Submitted          & \textbf{69.23}         & -                        \\
        \midrule[0.05em]
        No emoji           & 68.36                  & - 0.87                   \\
        \midrule[0.05em]
        No ELMo            & 65.52                  & - 3.71                   \\
        \midrule[0.05em]
        Concat Pooling     & 68.47                  & - 0.76                   \\
        \midrule[0.05em]
        LSTM hidden=4096   & 69.10                  & - 0.13                   \\
        LSTM hidden=1024   & 68.93                  & - 0.30                   \\
        LSTM hidden=512    & 68.43                  & - 0.80                   \\
        \midrule[0.05em]
        POS emb dim=100    & 68.99                  & - 0.24                   \\
        POS emb dim=75     & 68.61                  & - 0.62                   \\
        POS emb dim=50     & 69.33                  & + 0.10                   \\
        POS emb dim=25     & 69.21                  & - 0.02                   \\
        \midrule[0.05em]
        SGD optim lr=1     & 64.33                  & - 4.90                   \\
        SGD optim lr=0.1   & 66.11                  & - 3.12                   \\
        SGD optim lr=0.01  & 60.72                  & - 8.51                   \\
        SGD optim lr=0.001 & 30.49                  & - 38.74                  \\
        \bottomrule

@@

Using a greater BiLSTM hidden size did not help the model, probably because of
the reason mentioned earlier; the fully-connected layer was not big or deep enough
to exploit the additional information. Similarly, using a smaller hidden size
neither helped.

We found that using 50-dimensional part-of-speech embeddings slightly improved results,
which implies that better fine-tuning this hyperparameter, or using a better POS
tagger could yield an even better performance.

@@

    \caption{Dropout Ablation.}
    \vspace{-0.4cm}

@@

\subsection{Error Analysis}

Figure~\ref{fig:confusion_matrix} shows the confusion matrix of the model
evaluated on the test set, and Table~\ref{table:classification_report} the
corresponding classification report.

\begin{figure}[!h]
    \centering
    \includegraphics[width=\columnwidth]{images/confusion_matrix.pdf}

    \caption{Confusion Matrix (Test Set).}
    % \vspace{-0.4cm}

    % \caption*{
    %     \footnotesize Rows correspond to the dropout applied both after
    %     the ELMo layer (word encoding layer) and after the fully-connected network's
    %     hidden layer, while columns correspond to the dropout applied after the
    %     max-pooling operation (sentence encoding layer.)}

    \label{fig:confusion_matrix}
\end{figure}

\begin{table}[htpb]
    \centering
    \footnotesize

    \begin{tabular}{cccc}

        \toprule
                          & \textbf{Precision} & \textbf{Recall} & \textbf{F1-score} \\
        \midrule
        \texttt{anger}    & 0.643     & 0.601  & 0.621    \\
        \texttt{disgust}  & 0.703     & 0.661  & 0.682    \\
        \texttt{fear}     & 0.742     & 0.721  & 0.732    \\
        \texttt{joy}      & 0.762     & 0.805  & 0.783    \\
        \texttt{sad}      & 0.685     & 0.661  & 0.673    \\
        \texttt{surprise} & 0.627     & 0.705  & 0.663    \\
        \midrule
        Average           & 0.695     & 0.695  & 0.694    \\
        \bottomrule
    \end{tabular}

    \caption{Classification Report (Test Set).}

\label{table:classification_report}
\end{table}

@@

        \toprule
        \textbf{} & \textbf{Present} & \textbf{Not Present} \\
        \midrule
        Emoji     & 4805 (76.6\%)    & 23952 (68.0\%)       \\
        Hashtags  & 2122 (70.5\%)    & 26635 (69.4\%)       \\
        \bottomrule

@@

\begin{table}[!h]
    \centering
    \scriptsize
        \begin{tabular}{lrrcrcr}

            \toprule
           \multirow{2}{*}{\textbf{Emoji alias}} & \multirow{2}{*}{\textbf{N}\hspace{0.1cm}} & \multicolumn{2}{c}{\textbf{emoji}} & \multicolumn{2}{c}{\textbf{no-emoji}} &
            \multirow{2}{*}{$\bf{\Delta}$\textbf{\%}} \\
            \cmidrule(lr){3-4}
            \cmidrule(lr){5-6}
                                &     & \#   & \% & \# & \% & \\
            \midrule[0.08em]
            \texttt{mask}       & 163   \hspace{0.1cm} & 154 & 94.48 \hspace{0.1cm} & 134 & 82.21 & - 12.27 \\
          \texttt{two\_hearts}  & 87    \hspace{0.1cm} & 81  & 93.10 \hspace{0.1cm} & 77  & 88.51 & - 4.59  \\
          \texttt{heart\_eyes}  & 122   \hspace{0.1cm} & 109 & 89.34 \hspace{0.1cm} & 103 & 84.43 & - 4.91  \\
          \texttt{heart}        & 267   \hspace{0.1cm} & 237 & 88.76 \hspace{0.1cm} & 235 & 88.01 & - 0.75  \\
          \midrule[0.001em]
          \texttt{rage}         & 92    \hspace{0.1cm} & 78  & 84.78 \hspace{0.1cm} & 66  & 71.74 & - 13.04 \\
          \texttt{cry}          & 116   \hspace{0.1cm} & 97  & 83.62 \hspace{0.1cm} & 83  & 71.55 & - 12.07 \\
          \texttt{sob}          & 490   \hspace{0.1cm} & 363 & 74.08 \hspace{0.1cm} & 345 & 70.41 & - 3.67  \\
          \texttt{unamused}     & 167   \hspace{0.1cm} & 121 & 72.46 \hspace{0.1cm} & 116 & 69.46 & - 3.00  \\
          \midrule[0.001em]
          \texttt{weary}        & 204   \hspace{0.1cm} & 140 & 68.63 \hspace{0.1cm} & 139 & 68.14 & - 0.49  \\
          \texttt{joy}          & 978   \hspace{0.1cm} & 649 & 66.36 \hspace{0.1cm} & 629 & 64.31 & - 2.05  \\
          \texttt{sweat\_smile} & 111   \hspace{0.1cm} & 73  & 65.77 \hspace{0.1cm} & 75  & 67.57 & 1.80 \\
          \texttt{confused}     & 77    \hspace{0.1cm} & 46  & 59.74 \hspace{0.1cm} & 48  & 62.34 & 2.60 \\
          \bottomrule

@@

Table \ref{table:emoji_fine_grained} shows the effect specific emoji have on
classification performance. It is clear some emoji strongly contribute to improving prediction quality. The
most interesting ones are \texttt{mask}, \texttt{rage}, and \texttt{cry}, which
significantly increase accuracy. Further, contrary to intuition, the
\texttt{sob} emoji contributes less than \texttt{cry}, despite representing a
stronger emotion. This is probably due to \texttt{sob} being used for depicting
a wider spectrum of emotions.

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=\columnwidth]{images/acc_vs_tdp_variation.eps}
    \caption{Effect of the amount of training data on classification performance.}
    \label{fig:data_amt_vs_acc}
\end{figure}
